---
title: "LLM Deliberation Experiment"
subtitle: "`r params$pilot_id`"
output:
  ioslides_presentation:
    widescreen: true
    css: report_style.css
params:
  pilot_id: "test_full_001"
  project_root: ".."
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.width = 8,
  fig.height = 4.5,
  dpi = 100,
  out.width = "100%",
  fig.align = "center"
)

library(tidyverse)
library(jsonlite)
library(scales)

# Condition labels and colors
condition_labels <- c(
  "simple_voting" = "Simple Voting",
  "simple_passive" = "Simple Passive",
  "clarified_passive" = "Clarified Passive",
  "acp" = "ACP"
)

condition_colors <- c(
  "simple_voting" = "#CCCCCC",
  "simple_passive" = "#A6CEE3",
  "clarified_passive" = "#1F78B4",
  "acp" = "#6A3D9A"
)

# Custom theme - smaller for slides
theme_slides <- function() {
  theme_minimal(base_size = 11) +
    theme(
      plot.title = element_text(face = "bold", size = 14),
      plot.subtitle = element_text(color = "gray40", size = 10),
      legend.position = "bottom",
      panel.grid.minor = element_blank()
    )
}

# Load data
pilot_dir <- file.path(params$project_root, "outputs", params$pilot_id)
json_data <- fromJSON(file.path(pilot_dir, "participants.json"), simplifyDataFrame = FALSE)
participants_raw <- json_data$participants

# Parse participants
participants <- map_dfr(participants_raw, function(p) {
  tibble(
    participant_id = p$participant_id %||% NA_character_,
    condition = p$condition %||% NA_character_,
    status = p$status %||% NA_character_,
    initial_choice = p$initial_choice %||% NA_character_,
    final_choice = p$final_choice %||% NA_character_,
    initial_explanation = p$initial_explanation %||% NA_character_,
    age = p$demographics$age %||% NA_integer_,
    sex = p$demographics$sex %||% NA_character_,
    education = p$demographics$education %||% NA_character_,
    political_leaning = p$demographics$political_leaning %||% NA_character_,
    income = p$demographics$income %||% NA_character_,
    has_clarification = !is.null(p$clarification_transcript) && length(p$clarification_transcript) > 0,
    has_adversarial = !is.null(p$adversarial_transcript) && length(p$adversarial_transcript) > 0
  )
}) %>%
  mutate(
    position_changed = !is.na(initial_choice) & !is.na(final_choice) & initial_choice != final_choice
  )

# Load config if available
config_file <- file.path(pilot_dir, "config.yaml")
config <- if (file.exists(config_file)) {
  yaml::read_yaml(config_file)
} else {
  list(topic = list(description = "Unknown topic"))
}

# Summary stats
n_total <- nrow(participants)
n_complete <- sum(participants$status == "complete", na.rm = TRUE)
conditions <- unique(participants$condition)
```

## Experiment Overview

**Topic:** `r config$topic$description %||% "LLM-mediated collective decision making"`

**Pilot ID:** `r params$pilot_id`

<div class="columns-2">

**Participants:**

- Total: `r n_total`
- Complete: `r n_complete`
- Conditions: `r length(conditions)`

**Conditions:**

```{r overview-table, results='asis'}
participants %>%
  filter(status == "complete") %>%
  count(condition, name = "n") %>%
  mutate(condition = condition_labels[condition]) %>%
  knitr::kable(col.names = c("Condition", "N"), format = "html")
```

</div>

## Initial Vote Distribution

```{r initial-votes}
vote_counts <- participants %>%
  filter(status == "complete", !is.na(initial_choice)) %>%
  count(initial_choice, name = "n") %>%
  mutate(pct = n / sum(n))

ggplot(vote_counts, aes(x = reorder(initial_choice, -n), y = n, fill = initial_choice)) +
  geom_col(width = 0.7) +
  geom_text(aes(label = paste0(n, " (", percent(pct, accuracy = 1), ")")),
            vjust = -0.3, size = 3.5) +
  scale_fill_brewer(palette = "Set2", guide = "none") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Initial Vote Distribution",
    subtitle = "Before any deliberation or exposure",
    x = NULL,
    y = "Count"
  ) +
  theme_slides() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Position Change Rates {.flexbox .vcenter}

```{r change-rates}
change_data <- participants %>%
  filter(status == "complete", !is.na(position_changed)) %>%
  group_by(condition) %>%
  summarise(
    n = n(),
    changed = sum(position_changed),
    change_rate = mean(position_changed),
    se = sqrt(change_rate * (1 - change_rate) / n),
    .groups = "drop"
  ) %>%
  mutate(
    condition = factor(condition, levels = names(condition_labels))
  )

ggplot(change_data, aes(x = condition, y = change_rate, fill = condition)) +
  geom_col(width = 0.7) +
  geom_errorbar(
    aes(ymin = pmax(0, change_rate - 1.96 * se),
        ymax = pmin(1, change_rate + 1.96 * se)),
    width = 0.2, linewidth = 0.8
  ) +
  geom_text(
    aes(label = paste0(round(change_rate * 100, 1), "%\n(", changed, "/", n, ")")),
    vjust = -0.5, size = 3, fontface = "bold"
  ) +
  scale_fill_manual(values = condition_colors, guide = "none") +
  scale_x_discrete(labels = condition_labels) +
  scale_y_continuous(
    labels = percent_format(),
    limits = c(0, NA),
    expand = expansion(mult = c(0, 0.2))
  ) +
  labs(
    title = "Position Change Rate by Condition",
    subtitle = "Percentage of participants who changed their vote after deliberation",
    x = NULL,
    y = "Change Rate"
  ) +
  theme_slides()
```

## Vote Shifts (Sankey)

```{r sankey, fig.height=5}
# Check if ggalluvial is available
if (requireNamespace("ggalluvial", quietly = TRUE)) {
  library(ggalluvial)

  flow_data <- participants %>%
    filter(status == "complete", !is.na(initial_choice), !is.na(final_choice)) %>%
    count(condition, initial_choice, final_choice, name = "freq") %>%
    mutate(
      changed = initial_choice != final_choice,
      condition = factor(condition, levels = names(condition_labels))
    )

  if (nrow(flow_data) > 0) {
    ggplot(flow_data, aes(axis1 = initial_choice, axis2 = final_choice, y = freq)) +
      geom_alluvium(aes(fill = changed), width = 1/3, alpha = 0.7) +
      geom_stratum(width = 1/3, fill = "gray90", color = "gray40") +
      geom_text(stat = "stratum", aes(label = after_stat(stratum)), size = 1.8) +
      scale_x_discrete(limits = c("Initial", "Final"), expand = c(0.15, 0.05)) +
      scale_fill_manual(
        values = c("FALSE" = "gray70", "TRUE" = "#E41A1C"),
        labels = c("FALSE" = "No change", "TRUE" = "Changed"),
        name = NULL
      ) +
      facet_wrap(~condition, labeller = labeller(condition = condition_labels), ncol = 2) +
      labs(
        title = "Vote Flows: Initial to Final Position",
        subtitle = "Red flows indicate position changes"
      ) +
      theme_slides() +
      theme(
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        panel.grid = element_blank()
      )
  } else {
    ggplot() +
      annotate("text", x = 0.5, y = 0.5, label = "Insufficient data for Sankey diagram") +
      theme_void()
  }
} else {
  ggplot() +
    annotate("text", x = 0.5, y = 0.5, label = "ggalluvial package not available") +
    theme_void()
}
```

## Dialogue Engagement

```{r dialogue-engagement}
# Calculate engagement metrics from raw data
engagement_data <- map_dfr(participants_raw, function(p) {
  if (is.null(p$condition) || p$status != "complete") return(NULL)

  clar_words <- 0
  if (!is.null(p$clarification_transcript) && length(p$clarification_transcript) > 0) {
    clar_words <- sum(sapply(p$clarification_transcript, function(x) {
      str_count(x$content %||% "", "\\S+")
    }))
  }

  adv_words <- 0
  if (!is.null(p$adversarial_transcript) && length(p$adversarial_transcript) > 0) {
    adv_words <- sum(sapply(p$adversarial_transcript, function(x) {
      str_count(x$content %||% "", "\\S+")
    }))
  }

  tibble(
    participant_id = p$participant_id,
    condition = p$condition,
    clarification_words = clar_words,
    adversarial_words = adv_words,
    total_words = clar_words + adv_words
  )
})

if (nrow(engagement_data) > 0) {
  engagement_summary <- engagement_data %>%
    group_by(condition) %>%
    summarise(
      n = n(),
      mean_words = mean(total_words),
      se = sd(total_words) / sqrt(n()),
      .groups = "drop"
    ) %>%
    mutate(condition = factor(condition, levels = names(condition_labels)))

  ggplot(engagement_summary, aes(x = condition, y = mean_words, fill = condition)) +
    geom_col(width = 0.7) +
    geom_errorbar(
      aes(ymin = mean_words - 1.96 * se, ymax = mean_words + 1.96 * se),
      width = 0.2
    ) +
    geom_text(aes(label = round(mean_words)), vjust = -0.5, size = 3.5, fontface = "bold") +
    scale_fill_manual(values = condition_colors, guide = "none") +
    scale_x_discrete(labels = condition_labels) +
    scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
    labs(
      title = "Average Dialogue Word Count by Condition",
      subtitle = "Total words in clarification + adversarial dialogues",
      x = NULL,
      y = "Mean Word Count"
    ) +
    theme_slides()
} else {
  ggplot() +
    annotate("text", x = 0.5, y = 0.5, label = "No dialogue data available") +
    theme_void()
}
```

## Example: Position Changer

```{r example-dialogue, results='asis'}
# Find a position changer from ACP condition
changers <- participants %>%
  filter(condition == "acp", status == "complete", position_changed == TRUE)

if (nrow(changers) > 0) {
  example_id <- changers$participant_id[1]
  example_raw <- participants_raw[[which(sapply(participants_raw, function(p) p$participant_id == example_id))]]

  cat(paste0("**Participant:** ", example_id, "\n\n"))
  cat(paste0("**Initial Vote:** ", example_raw$initial_choice, "\n\n"))
  cat(paste0("**Final Vote:** ", example_raw$final_choice, " (CHANGED)\n\n"))
  cat("---\n\n")

  # Show adversarial dialogue excerpt
  if (!is.null(example_raw$adversarial_transcript) && length(example_raw$adversarial_transcript) > 0) {
    cat("**Adversarial Dialogue Excerpt:**\n\n")
    cat('<div class="dialogue-excerpt">\n\n')
    transcript <- example_raw$adversarial_transcript
    for (i in seq_len(min(4, length(transcript)))) {
      turn <- transcript[[i]]
      role <- ifelse(turn$role == "moderator", "Moderator", "Participant")
      content <- str_trunc(turn$content, 300)
      cat(paste0("> **", role, ":** ", content, "\n\n"))
    }
    cat('</div>\n\n')
  }
} else {
  cat("*No position changers found in ACP condition.*\n\n")

  # Show any ACP participant instead
  acp_participants <- participants %>% filter(condition == "acp", status == "complete")
  if (nrow(acp_participants) > 0) {
    example_id <- acp_participants$participant_id[1]
    example_raw <- participants_raw[[which(sapply(participants_raw, function(p) p$participant_id == example_id))]]

    cat(paste0("**Example ACP Participant:** ", example_id, "\n\n"))
    cat(paste0("**Vote:** ", example_raw$initial_choice, " (unchanged)\n\n"))
  }
}
```

```{r full-dialogue, results='asis'}
# Show complete adversarial transcript split across multiple slides
if (exists("example_raw") && !is.null(example_raw$adversarial_transcript) && length(example_raw$adversarial_transcript) > 0) {
  transcript <- example_raw$adversarial_transcript

  # Calculate how many turns per slide based on content length
  # Target ~800 chars per slide to fit comfortably with small font
  chars_per_slide <- 800

  # Group turns into slides
  slides <- list()
  current_slide <- list()
  current_chars <- 0

  for (i in seq_along(transcript)) {
    turn <- transcript[[i]]
    turn_chars <- nchar(turn$content)

    # Start new slide if this turn would exceed limit (but always include at least 1 turn per slide)
    if (current_chars + turn_chars > chars_per_slide && length(current_slide) > 0) {
      slides <- c(slides, list(current_slide))
      current_slide <- list(turn)
      current_chars <- turn_chars
    } else {
      current_slide <- c(current_slide, list(turn))
      current_chars <- current_chars + turn_chars
    }
  }
  # Add final slide
  if (length(current_slide) > 0) {
    slides <- c(slides, list(current_slide))
  }

  n_slides <- length(slides)

  # Output each slide
  for (s in seq_along(slides)) {
    if (n_slides > 1) {
      cat(paste0("\n\n## Full Dialogue: Position Changer (", s, "/", n_slides, ") {.smaller}\n\n"))
    } else {
      cat("\n\n## Full Dialogue: Position Changer {.smaller}\n\n")
    }

    # Header on first slide only
    if (s == 1) {
      cat(paste0("**Participant:** ", example_raw$participant_id, " | "))
      cat(paste0("**", example_raw$initial_choice, "** → **", example_raw$final_choice, "**\n\n"))
    }

    cat('<div class="dialogue-excerpt">\n\n')
    for (turn in slides[[s]]) {
      role <- ifelse(turn$role == "moderator", "Moderator", "Participant")
      cat(paste0("> **", role, ":** ", turn$content, "\n\n"))
    }
    cat('</div>\n')
  }
} else {
  cat("\n\n## Full Dialogue: Position Changer {.smaller}\n\n")
  cat("*No adversarial dialogue available for this participant.*\n")
}
```

## Key Takeaways

```{r takeaways, results='asis'}
# Calculate key stats
total_complete <- sum(participants$status == "complete")
total_changed <- sum(participants$position_changed, na.rm = TRUE)
overall_change_rate <- mean(participants$position_changed, na.rm = TRUE)

acp_data <- participants %>% filter(condition == "acp", status == "complete")
acp_change_rate <- mean(acp_data$position_changed, na.rm = TRUE)

baseline_data <- participants %>% filter(condition == "simple_passive", status == "complete")
baseline_change_rate <- mean(baseline_data$position_changed, na.rm = TRUE)

cat("### Summary Statistics\n\n")
cat(paste0("- **Total Participants:** ", total_complete, " completed the experiment\n"))
cat(paste0("- **Overall Change Rate:** ", round(overall_change_rate * 100, 1), "% changed their position\n"))
cat(paste0("- **ACP Change Rate:** ", round(acp_change_rate * 100, 1), "%\n"))
cat(paste0("- **Baseline (Simple Passive):** ", round(baseline_change_rate * 100, 1), "%\n\n"))

if (acp_change_rate > baseline_change_rate) {
  diff <- acp_change_rate - baseline_change_rate
  cat(paste0("### Main Finding\n\n"))
  cat(paste0("**ACP condition shows ", round(diff * 100, 1), " percentage points higher change rate**\n"))
  cat("than baseline passive exposure.\n")
} else {
  cat("### Main Finding\n\n")
  cat("No clear difference between ACP and baseline conditions observed.\n")
}
```

---

<div class="centered">
**Report generated:** `r Sys.time()`

**Pilot:** `r params$pilot_id`
</div>
